### 2011-08-06 paper.tex
revise Laura's first draft of paper

\begin{abstract}
        % Do not use inserted blank lines (ie \\) until main body of text.
        \paragraph*{Background:} Text for this section of the abstract.

        \paragraph*{Methods:} Text for this section of the abstract \ldots

        \paragraph*{Results:} Text for this section of the abstract \ldots

        \paragraph*{Conclusions:} Text for this section of the abstract \ldots
\end{abstract}

@@

\section*{Background}
  (FIXME: need to talk about GBD here...different models for each cause...different levels in the
  cause hierarchy).  Models for each cause are estimated separately, so there is no guarantee of 
  consistency and the estimated cause fractions for any given country-sex-year do not necessarily 
  sum to one. \pb

  (FIXME: this would be where a review of the literature would be appropriate, had we done one.) \pb

  A simulation environment was constructed to replicate important characteristics of 
  the problem described.  We propose two models that take different approaches to predicting 
  consistent cause fractions from inconsistent preliminary estimates.  These two models are then  
  validated in the simulation environment and compared on a variety of metrics. \pb

@@

\section*{Methods}
  \subsection*{Simulation Environment}
     Simulated cause fraction estimates are generated for 3 causes over a period of 4 years. 
    (FIXME: the number of causes and the length of the time period could change). 'True' 
    cause fractions are specified for each of the causes in each of the years. Additionally, 
    a 'true' precision is specified for each cause fraction, and this is treated as 
    constant over time (FIXME: should probably explain what this precision is supposed to 
    be telling us about the cause fractions). A single draw from a normal distribution 
    with mean 0 and precision equal to the true cause fraction precision is generated. 
    This is added to the true cause fractions to imitate an unbiased preliminary estimate of the
    true cause fractions. A set of 1000 draws from a normal distribution using this estimated
    cause fraction as the mean and the true precision as the precision is taken as the
    'data' for this simulation that will eventually be used in the modeling process. 
    This procedure is repeated, resulting in 50 sets of simulated preliminary 
    estimates for each specified set of cause fractions. All draws from normal distributions 
    are carried out in logit space to guarantee that the resulting cause fractions are each
    between zero and one. Both of the models described below are fit on each of the 50 sets 
    of preliminary estimates for each specified cause fraction scenario. \pb

    A variety of different scenarios are tested (FIXME: must decide on what the scenarios are
    before I can describe them) \pb

  \subsection*{Statistical Model}
    Two models are proposed for estimating consistent cause fractions. The first takes an obvious
    approach: all cause fractions are simply scaled by a factor equal to their sum: 
    (FIXME: need to put some math here, i.e. the bad model)
  \begin{align*}
    math!
  \end{align*}
    For the simulated estimates described above, this is carried out individually for 
    each simulation and then the mean estimate for each cause is taken as the final 
    estimate for each cause fraction. The 95\% high probability density region (95\% hpd) 
    is also calculated and used to generate estimates of the upper and lower bounds of 
    the final cause fraction estimates. \pb

    The second model proposed attempts to take into account differences in the estimated 
    precision of the preliminary cause fraction estimates. Additionally, it explicitly 
    takes into account correlation in these preliminary estimates over time (FIXME: this 
    may not actually be true. Must understand the model better before writing about it). 
    (FIXME: probably need a more thorough verbal description here of the model; also 
    need to update the formula Abie typed below). 
  \begin{align*}
    \Pr[Y | \pi, \tau] &= \sum_{i=1}^N \frac{1}{N} \Pr[Y_i | \pi, \tau]\\
    Y_i | \pi, \tau &\sim N(\pi, \tau)\\
    \pi &\sim \Dirichlet(1) \\
    \tau &\sim \Uniform[0, \infty) \\
  \end{align*}
    (FIXME: need to talk about how the model is fit). The mean of pi across all retained 
    iterations of the MCMC chain is taken as the final estimate of the cause fractions 
    and the 95\% hpd is calculated from all retained iterations. \pb

  \subsection*{Quality Metrics}
    The final cause fraction estimates from each model are compared to the known true cause 
    fractions and the absolute error, relative error, and coverage (i.e. the true estimate
    falls between the upper and lower bounds) are calculated for each cause and each 
    of the 50 sets of estimates. The mean and median absolute error, mean and 
    median relative error, and mean coverage are estimated by cause across all sets of 
    estimates. Mean coverage is also estimated for all causes and the 
    percent of times across all 50 sets of estimates that all causes are covered (full 
    coverage) is also computed. One additional metric, cause specific mortality fraction 
    accuracy (CSMF accuracy) is also computed for each of the 50 sets of estimates 
    according to the following formula: 
  \begin{align*}
    1-\sum \left | CSMF_{pred} - CSMF_{true} \right |/(2*(1-min(cf_{true})))
  \end{align*}
    The mean and median CSMF accuracy are then computed across all sets of estimates. \pb

    (FIXME: need to remove some of the metrics here that we aren't ultimately going to 
    use and also need to describe why each one of the ones that remain are interesting 
    and necessary given the others) \pb

>>>>

\begin{abstract}
        % Do not use inserted blank lines (ie \\) until main body of text.
        \paragraph*{Background:} Population-level information on cause-specific mortality rates by country, age, sex, and time is important for decision-making.  Models are crucial to estimating and predicting this information when direct measurements are not available.  State-of-the-art models produce estimates disease by disease, however, and do not enforce the constraint that the sum of deaths over every cause must be equal to the total deaths due to all causes.

        \paragraph*{Methods:} We developed a statistical model to correct for inconsistencies between cause-specific mortality rate estimates and all-cause mortality estimates. We compared our model to a naive approach in a validation environment using a series of simulation studies designed to test TK.  We measured the quality of the corrected estimates by TK.

        \paragraph*{Results:} Over all simulations, our approach performed TK, compared to TK for the naive approach.  For setting TK, our approach showed significant benefits in terms of TK and TK.  For setting TK, neither approach performed very well, demonstrating that without sufficiently accurate input data, all model-based approaches have limitations.

        \paragraph*{Conclusions:} Our model is a fast and effective way to improve the estimates produced by many separate models, and we recommend using it to enforce consistency between cause-specific mortality rates and all-cause mortality rates.
\end{abstract}

@@

\section*{Background}
  TK one paragraph on population health metrics, GBD, CoD paper which
  uses different ensemble models for each cause and how hard it would
  be to combine them.

  Models for each cause are estimated separately, so there is no
  guarantee of consistency and the estimated cause fractions for any
  given country-sex-year do not necessarily sum to one.

  TK one paragraph on what we have done, a model-based approach to
  enforcing consistency between cause-specific model estimates and
  all-cause estimates.  We validated our approach through a simulation
  study, with an environment constructed to capture important
  characteristics of the real challenge.  We propose two models
  that take different approaches to predicting consistent cause
  fractions from inconsistent preliminary estimates.  These two models
  are then validated in the simulation environment and compared on a
  variety of metrics.

  TK one paragraph on the importance of what we have done.

@@

\section*{Methods}
The estimation challenge we faced was the following: For a given sex
and geographic region, we were given estimates of the cause-specific
mortality fraction $F_j$ for a mutually exclusive, collectively
exhaustive list causes $j = 1,\ldots,J$.  Each estimate is
represented by $T$ draws from the posterior distribution over values
for $N$ years:
\[F_j = \left(f_{j,y_1}^t, \ldots, f_{j,y_2}^t\right)_{t=1}^T.
\]
Our goal was to produce a posterior distribution for the true cause
fractions $\pi_{j,y}$, by incorporating the consistency condition
$\sum_{j} \pi_{j,y}=1$ for all $y$.  This would allow us to produce
point estimates and uncertainty intervals for individual cause
fractions, as well as estimate the change in $\pi_{j,y}$ as a function
of $y$ for each $j$.

  \subsection*{Quality Metrics}
  We wanted the estimated cause fractions from any model to be close
  to the true cause fractions and the absolute error, although being
  close in relative error is also important for certain applications.
  Following the approach taken in Verbal Autopsy quality metrics
  \ref{VAMetrics}, we summarize the absolute error using cause
  specific mortality fraction accuracy (CSMF accuracy), defined by the
  following formula:
  \begin{align*}
    \text{CSMF Accuracy} = 1- \sum \frac{\left | \pi_{j,y} - \pi^{\text{true}}_{j,y} \right |}{2(1-\min_j(\pi^{\text{true}}_{j,y}))}
  \end{align*}
  Furthermore, we desired that the coverage probability (the
  probability that the true estimate falls between the upper and lower
  bounds of the $95\%$ uncertainty interval) for each cause be close
  to $0.95$.

  \subsection*{Statistical Models}
  Before developing a sophisticated statistical model for estimating
  $\pi_{j,y}$, we considered a naive approach: for each
  $t=1,\ldots,T$, for each $y$, we simply scaled the vector
  $(f_{1,y}^t, \ldots, f_{J,y}^t)$ to make it sum to unity:
  \begin{align*}
    \pi_{j,y}^t = \frac{f_{j,y}^t}{\sum_{j'=1}^J f_{j',y}^t}.
  \end{align*}
  This provides a non-parametric posterior distribution for
  $\pi_{j,y}$, which we summarized as a point estimate for each $j$
  and $y$ by taking the mean over values of $t$, and estimates
  uncertainty intervals by taking the $95\%$ highest probability
  density interval.

  We developed the more traditional statistical model that we now
  describe to take into account the differences in uncertainty that
  the different $f_{j,y}$ in the input data often exhibit.  The formal
  specification of the model is the following:
  \begin{align*}
    p(f_{j,y}^1,\ldots,f_{j,y}^T | \pi) &\propto \exp\left\{-\frac{\left(\pi_{j,y}-\E_t[f_{j,y}]\right)^2}{\Var_t[f_{j,y}]} \right\}\\
    \pi_{j,y} &= \frac{\alpha_{j,y}^t}{\sum_{j'=1}^J \alpha_{j',y}^t}\\
    \alpha_{j,y} &\sim \Normal\left(0, 1^2\right)
  \end{align*}

 This approach can be modified explicitly account for time-correlation in the
 $f_{j,y}$ inputs, by making the likelihood look more like a
 multivariate normal:
  \begin{align*}
    p\left((f_{j,y_1}^t,\ldots,f_{j,y_N}^t)_{t=1}^T \big| \pi \right)
    &\propto \exp\left\{-\left((\pi_j - \E_t[f_j])^T \Sigma^{-1} (\pi_j - \E_t[f_j])\right) \right\}\\
    \pi_{j,y} &= \frac{\alpha_{j,y}^t}{\sum_{j'=1}^J \alpha_{j',y}^t}\\
    \alpha_{j,y} &\sim \Normal\left(0, 1^2\right)\\
    \Sigma &= \Matern(\sigma, \rho, 2) + \Diag(\Var_t[f_{j,y_1}],
      \ldots, \Var_t[f_{j,y_N}])\\
    \sigma &\sim \Uniform[0,1]\\
    \rho &\sim \Normal_{\epsilon+}(20, 10^2)
  \end{align*}

  \subsection*{Simulation Environment}
  To validate the models above, compare the sophisticated approaches
  to the naive approach, and generally assess how much improvement
  this sort of consistenty constraint can be expected to provide, we
  conducted a simulation study.  We chose true cause fractions for a
  three different sized cause lists, for a 10 year
  time period, and generated noisy ``estimates'' of the true cause
  fractions to produce simulated input data for which ground truth was
  known.  We varied the dispersion of the estimates, as well as the
  bias to understand how the different models perform in different
  settings.  This can be understood as a three-armed factorial study
  design, with cause lists of size $J=3, 30, 300$, dispersion taking
  values low, medium, and high, and bias taking values low, medium,
  and high.

  To be precise the true mean cause fractions were selected random from an
  uninformative Dirichlet distribution with $J$ dimensions. These were
  then extended to smoothly varying functions of time by TK. Then bias
  was generated by taking a single draw from a normal distribution
  with mean 0 and variance $\sigma_b^2$, which was added to the true
  cause fractions in logit space to introduce systematic bias. We then
  generated a collection of 1000
  draws from multivariate normal distributions using the logit of time-series
  of the biased cause fraction
  as the mean and covariance function $\Matern(\sigma_d, 20,
  2)$. The inverse logit of these $1000$ samples (of $J\times 10$-dimensional vectors)
  constitute the simulated input data we used to validate our models.

  We repeated this procedure $500$ times for each of the $27$
  possibilities in the design, and calculated CSMF accuracy, coverage
  probability, and TK something about change over time for all of
  them.
